Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2019-10-09 tables/stormseverity.tex
Work on paper, use 34 knot threshold

% latex table generated in R 3.4.3 by xtable 1.8-2 package
% Wed Oct 10 12:11:35 2018
\begin{table}[ht]
\centering
\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.     
	 The median and interquartile range of number of county exposures per tropical cyclone are based on
         the tropical cyclones for which at least one US county was exposed. The years for which
         data are available for each metric are given in Table 1.} 
\label{tab:exposuresummaries}
\begin{tabular}{p{1.75cm}p{3cm}p{4cm}p{4cm}}
  \toprule
Metric & Mean (interquartile range) of county exposures per year & Median (interquartile range) of county exposures per tropical cyclone & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
  \midrule
Distance & 433 (224, 503) & 84 (23, 173) & Beryl, 1994 (330) \\ 
  Rain & 401 (152, 553) & 68 (14, 146) & Frances, 2004 (464) \\ 
  Wind & 224 (83, 353) & 34 (8, 80) & Ike, 2008 (355) \\ 
  Flood & 213 (84, 249) & 26 (6, 57) & Ivan, 2004 (317) \\ 
  Tornado & 53 (16, 43) & 7 (2, 16) & Ivan, 2004 (91) \\ 
   \bottomrule
\end{tabular}
\end{table}

>>>>

% latex table generated in R 3.6.1 by xtable 1.8-4 package
% Tue Oct  8 14:03:24 2019
\begin{table}[ht]
\centering
\caption{Summary statistics for the number of county tropical cyclone exposures under each metric.     
	 The median and interquartile range of number of county exposures per tropical cyclone are based on
         the tropical cyclones for which at least one US county was exposed. The years for which
         data are available for each metric are given in Table 1.} 
\label{tab:exposuresummaries}
\begin{tabular}{p{1.75cm}p{3cm}p{4cm}p{4cm}}
  \toprule
Metric & Mean (interquartile range) of county exposures per year & Median (interquartile range) of county exposures per tropical cyclone & Tropical cyclone with most counties exposures (\# exposed counties) \\ 
  \midrule
Rain & 401 (152, 553) & 68 (14, 146) & Frances, 2004 (464) \\ 
  Wind & 143 (51, 232) & 25 (14, 58) & Ike, 2008 (251) \\ 
  Flood & 213 (84, 248) & 26 (6, 58) & Ivan, 2004 (318) \\ 
  Tornado & 53 (16, 43) & 7 (2, 16) & Ivan, 2004 (91) \\ 
   \bottomrule
\end{tabular}
\end{table}

### 2019-10-09 writing/darlings_discussion.tex
Work on paper, use 34 knot threshold

\section*{Discussion}

Tropical cyclones impact public health in many ways, and epidemiological
studies of the health risk and impacts associated with tropical cyclone
exposures could help improve preparedness for and response to future tropical
cyclones.  However, tropical cyclones are multi-hazard events, making it
complicated to measure exposure. Here, we provide open-source county-level data
on multiple storm exposure metrics, and we find that county-level tropical
cyclone exposure assessments vary substantially when using a distance-based
metric versus four hazard-based metrics, as well as among different
hazard-based metrics. Our results can inform exposure assessment for future
county-level studies of the health risks and impacts associated with tropical
cyclones exposures, and the open-source data and software we have created
provides multi-hazard, county-level exposure data for such studies.

\subsection*{Exposure assessment data and software}

In creating this hurricane exposure dataset, we aimed for data that are
available for all eastern \ac{US} counties. In addition to providing tropical
cyclone exposure metrics for individual hazards, this data and its associated
software allow users not only to access measurements for single hazards, but
also to create tropical cyclone exposure profiles based on multiple hazards or
craft exposure indices that combine hazard
metrics~\parencite{chakraborty2005population, peduzzi2009assessing}. This
ability can be critical, as different hazards of tropical cyclones often act
synergistically in causing impacts~\parencite{smith2009}.

\begin{comment}
To assist with future tropical cyclone studies, we created and published open
source software with this dataset and accompanying
tools~\parencite{hurricaneexposuredata, hurricaneexposure}.  Many previous
studies have used geographical information system software (e.g., ArcGIS) to
assess exposure to tropical cyclones in the \ac{US}~\parencite{grabich2016,
zandbergen2009, czajkowski2011, kruk2010}.  Here, we offer methods to map and
output historic exposure to tropical cyclones that does not require the use of
proprietary software but instead uses a package written in the R statistical
programming language~\parencite{R}, which is free and open-source.  Further, by
including measurements of different hazard exposures in each county for each
tropical cyclone, this software allows for the development of more complex
exposure indices or models (e.g., random forests, multivariable generalized
linear models) that incorporate multiple tropical cyclone hazard measurements.  
\end{comment}

We investigated how well these data correspond with data from other available
sources (Figures~\ref{fig:raincomparison}\,--\,\ref{fig:floodcomparison}
and~S2). While generally in agreement with data from other sources,
there are a few caveats that should be considered when applying these data in
future studies. The rainfall data are from re-analysis data are generally
well-correlated with observed ground-based station data, but may sometimes
oversmooth very high rainfall values compared to ground-based observations
(Figure~\ref{fig:raincomparison}). When rainfall data is used to create binary
exposure classifications, this disagreement is unlikely to influence results,
as both data sources agree in identifying these as storms with high rainfall,
but it may be important to consider for cases that include rainfall as a
continuous measurement.

The wind data are based on modeled, rather than observed, values, and while the
modeled wind data generally agree well with post-analysis maximum wind radii
from \ac{HURDAT2}~\parencite{landsea2013} (Figure~\ref{fig:windcomparison}),
there were a few storms where the two datasets did not agree as well ($<$90\%
of counties classified into the same wind categories by the two sources of wind
estimates). These storms were exceptionally large systems for which high winds
persisted well inland from landfall, like Hurricanes Sandy and Ike
(Figures~\ref{fig:windcomparison} and~S2). Since the wind radii are based on
\ldots, they have [advantages]; their main disadvantages is that they provide
categorical, rather than continuous, estimates of wind speed, and in many cases
researchers might prefer more precise estimates of storm wind speeds. However,
since the wind radii data may be useful in some cases as either a primary
metric or for sensitivity analysis, we have include county-level wind estimates
for each storm based on these wind radii, in addition to the continuous modeled
windspeeds, in the open-source data \parencite{hurricaneexposuredata}.

For the flooding events data, we found it often corresponded with measurements
from \ac{US} Geographical Survey streamgages, with flood event listings more
likely during storms which elevated stream flows
(Figure~\ref{fig:floodcomparison}. However, there are differences between the
two flooding datasets, and these highlight both the difficulty of measuring
flood exposure at the county level and inherent challenges in using the storm
event database for epidemiologic exposure assessment.  The flood and tornado
data came from the \ac{NOAA} Storm Events database, which, while a widely-used
database of events maintained by \ac{NOAA}, is based on reports, and so may be
prone to underreporting~\parencite{Ashley2008flood, Curran2000}, especially in
less populated areas~\parencite{Witt1998, Ashley2007}, as well as to other
reporting errors.

\begin{comment}
[For discussion?] Based on these results, the two data sources were generally
in agreement, but the analysis highlights the difficulty of assessing
storm-related flooding at a multi-county, multi-year scale, as flooding can be
very localized within a county or can \ldots .   
\end{comment}

While these are important caveats for the data, we selected these data sources
as among the best currently available for measuring each of these hazards at a
multi-county, multi-year scale.

\subsection*{Patterns in tropical cyclone exposures}

We found geographic patterns in tropical cyclones based on the different
exposure metrics (Figure~\ref{fig:averageexposures}). These patterns were not
unexpected based on what is known about tropical cyclone hazards, but still
highlight variations across metrics that are critical to consider in designing
studies for tropical cyclone epidemiology, and they demonstrate the need for
multi-hazard exposure datasets for tropical cyclone epidemiology. Wind-based
exposures had a strong coastal pattern, which is consistent with the dramatic
decrease in wind intensity that typically characterizes the landfall of
tropical cyclones. Exposures by rain- and distance-based metrics often extended
further inland compared to the wind-based metric, up to the Appalachian
mountains. This agrees with previous research indicating that the Appalachian
mountains' topography both enhances precipitation during tropical cyclones and
provides hydrological conditions for severe flooding~\parencite{rees2001}.
Almost all tornado-based exposures were in southern coastal states, echoing
previous findings that most noteworthy tropical cyclone-related tornadoes occur
on the right-side of the tropical cyclone track in Atlantic-basin \ac{US}
storms~\parencite{moore2012}.  These exposure averages are somewhat limited as
estimates of long-term frequencies, as tropical cyclones follow decadal
patterns~\parencite{kossin2007more} likely not adequately captured in the
available data. However, exposure assessments spanning over a century have
found similar patterns for distance- and wind-based
metrics~\textcite{zandbergen2009, kruk2010} [double-check that refs found
similar patterns].

\begin{comment}
However, these frequency maps, together with evidence from
specific tropical cyclones (Figures~\ref{fig:ivanexposure}
and~\ref{fig:jaccard}),  do illustrate the potential for strong differences in
spatial patterns in tropical cyclone exposures, depending on which tropical
cyclone hazards are considered.  A few previous studies have sought to
determine county-level exposure to tropical cyclones over multi-year periods,
including~\textcite{zandbergen2009}, which estimated exposure in \ac{US}
counties to all \ac{US} landfalling Atlantic-basin tropical cyclones
between~1851 and~2003, using both a distance-based metric and a metric that
combined distance and windspeed, and~\textcite{kruk2010}, which explored
exposure to hurricane-related winds in the \ac{US}, including inland areas,
for~1900\,--\,2008.  Our results suggest that such exposure assessments may
perform well in capturing some tropical cyclone hazards (e.g., wind), but
likely miss other potentially dangerous tropical cyclone exposures, especially
for hazards that repeatedly threaten northern or inland counties (e.g., rain,
flooding).
\end{comment}

\subsection*{Agreement between exposure metrics}

We found that there was typically only moderate or low agreement between
county-level exposure classifications for a tropical cyclone when comparing
different exposure metrics
(Figures~\ref{fig:ivanexposure}\,--\,\ref{fig:jaccard}  and S4\,--\,S5). This
suggests that the use of one of these metrics as a proxy for all storm-based
hazards in epidemiological research would often be problematic, introducing
exposure misclassification.

\begin{comment}
For example, distance from a tropical cyclone's track is relatively easy to
measure and has been used as an operational metric of exposure to tropical
cyclones in previous large-scale studies (examples include ...). Since distance
itself does not constitute a hazard, distance is meant in these cases as a
surrogate to capture exposure to hazards from the tropical cyclone. However,
here we found that in assessing \ac{US} county-level exposure to tropical
cyclones, distance is, at best, a moderate, and often a very poor, surrogate
for exposure to the specific tropical cyclone hazards of high wind, extreme
rainfall, flooding, and tornadoes (Figure~\ref{fig:jaccard}). Therefore, use of
distance to assess tropical cyclone exposure for impact studies could result in
problematic exposure misclassification, which could mask true associations,
even strong associations, between tropical cyclone exposure and outcomes of
interest in impact studies~\parencite{savitz2016interpreting,
armstrong1998effect}.  
\end{comment}

These findings align with previous results from atmospheric science and related
fields on the characteristics of tropical cyclones. While tropical cyclone
rainfall and windspeed are often well-correlated when the tropical cyclone is
over water~\parencite{cerveny2000}, this relationship often does not remain as
strong once the hurricane has made landfall~\parencite{jiang2008}.  Fast-moving
tropical cyclones bring higher risks of dangerous winds
inland~\parencite{kruk2010}, while slow-moving tropical cyclone are likely to
bring more rain~\parencite{rappaport2000} and cause more damage because of
sustained hazardous conditions~\parencite{rezapour2014}. Further, while the
likelihood and extent of flooding during a tropical cyclone is related to the
tropical cyclone's rainfall, it is also driven by factors like top soil
saturation and the structure of the water basin's drainage
network~\parencite{chen2015, rees2001}.

However, we did find a small set of tropical cyclones for which for which
agreement between metrics were unusually high (e.g., Floyd in~1999, Irene in
2011, Hannah in~2008, Bertha in~1996; Figure~\ref{fig:jaccard}).  Hurricanes
Floyd in~1999 (Figure~S4) and Irene in~2011 both made their first \ac{US}
landfall in North Carolina at minor hurricane windspeeds (Category~2 and~1,
respectively) and then skimmed the eastern coast of the \ac{US} north through
New England, bringing large rainfall to much of the eastern coast from North
Carolina north and causing extensive inland flooding in North Carolina (Floyd)
and New England (Irene)~\parencite{avila2013atlantic, lawrence2000atlantic}.
Hurricanes Hannah in~2008 and Bertha in~1996 followed the eastern coastline
north of North Carolina in a pattern similar to Floyd and Irene. For this set
of storms, the tropical cyclones' persistent proximity to water may have helped
maintain wind speeds in similar patterns to rain and distance exposures,
resulting in this moderate to good agreement among exposure assessments based
on different metrics.

In some cases, exposure misclassification from using one of these exposure
metrics to capture one or more of the other storm hazards may be differential
(i.e., associated with the outcome of interest or with factors associated with
risk of the outcome of interest). For example, tropical cyclone wind exposures
tend to be concentrated in counties near the coast, since most tropical
cyclones rapidly decrease in their sustained windspeed following landfall.
However, tropical cyclone exposures based on distance can extend well inland,
following the tropical cyclone's tracks, but may not adequately capture all
wind-exposed counties near the coast. In this case, if the
etiologically-relevant exposure is high wind but exposure is classified based
on distance, the probability of being misclassified as unexposed would be
higher in coastal counties, while the probability of being misclassified as
exposed would be higher in inland counties. If coastal counties differ from
inland counties in either the outcome of interest or in factors associated with
risk of that outcome, differential exposure misclassification would
exist~\parencite{savitz2016interpreting}. Such differential exposure
misclassification can bias estimates of tropical cyclone effects either towards
the null (estimating a lower or null association compared to the true
association that exists) or away from the null (estimating a larger association
than actually exists)~\parencite{savitz2016interpreting, armstrong1998effect}.

\begin{comment}
The use of a single hazard-based metric (e.g., wind) could cause similar
problems if the impact is driven, at least in part, by a different hazard or by
multiple hazards of the tropical cyclone.
\end{comment}

\subsection*{Case study}

The conducted case study demonstrates how differences in exposure assessment
across storm exposure metrics can result in differing estimates of expected
physical exposure to tropical cyclones among a population
(Figure~\ref{fig:topelecdependexposure}).  Previous assessments of physical
exposure to tropical cyclones have often focused on solely storm winds and
storm proximity. For example, a global study of exposure and vulnerability to
natural hazards used a combination of windspeed above a certain threshold and
distance from the storm's track to assess
exposure~\parencite{peduzzi2009assessing}. While many storm impacts might be
most strongly linked to wind hazards, there is a growing recognition of the
potential risks of adverse health outcomes and property damage from rain- and
flood-related tropical cyclone hazards in the \ac{US}~\parencite{smith2009}.
Such assessments may underemphasize potentially dangerous exposures to certain
regions of the \ac{US}, particularly inland locations. In fact, based on our
assessment of physical exposures to tropical cyclones among
electricity-dependent Medicare beneficiaries, the highest expected physical
exposure under any metric was measured for Philadelphia County, PA, under the
flood metric (Figure~\ref{fig:topelecdependexposure}).  While this county has a
similar-sized population of electricity-dependent Medicare beneficiaries as
Miami-Dade County, FL (Figure~\ref{fig:topelecdependexposure}), the frequency
of storm-related flood events in this inland county results in a very high
expected physical exposure among this subpopulation, while this county is not
included in the top ten counties for physical exposure based on most of the
other exposure metrics.

\subsection*{Conclusions}

Previous research has highlighted the range of impacts that tropical cyclones
can have in \ac{US} communities. However, studies of tropical cyclone exposures
and their impacts have varied widely in how they assess exposure. Here we found
large differences in which counties are exposed to different hazards of
tropical cyclones and that distance is, at best, a moderate, and often a very
poor, surrogate for exposure to the specific tropical cyclone hazards of high
wind, extreme rainfall, flooding, and tornadoes. Use of distance as a surrogate
for any of these hazards could lead to exposure misclassification and, in the
case of tropical cyclone risk and impact studies, including epidemiological
studies, result in biased estimates.  Similarly, if such studies disagree on
their findings, it could result from the often poor agreement between exposure
classifications based on these different metrics, making studies that use
different metrics hard to compare or combine in meta-analyses. Our findings
highlight the importance of clarifying the potential pathway from tropical
cyclone hazards to health impacts when conducting tropical cyclone
epidemiological studies, and then basing exposure assessment on these hazards.
To help, we provide open-source tropical cyclone hazard-related measurements
for a large collection of historical Atlantic-basin cyclones, aggregated at the
county level to align with the spatial scale at which much health-related data
is collected.

>>>>

To assist with future tropical cyclone studies, we created and published open
source software with this dataset and accompanying
tools~\parencite{hurricaneexposuredata, hurricaneexposure}.  Many previous
studies have used geographical information system software (e.g., ArcGIS) to
assess exposure to tropical cyclones in the \ac{US}~\parencite{grabich2016,
zandbergen2009, czajkowski2011, kruk2010}.  Here, we offer methods to map and
output historic exposure to tropical cyclones that does not require the use of
proprietary software but instead uses a package written in the R statistical
programming language~\parencite{R}, which is free and open-source.  Further, by
including measurements of different hazard exposures in each county for each
tropical cyclone, this software allows for the development of more complex
exposure indices or models (e.g., random forests, multivariable generalized
linear models) that incorporate multiple tropical cyclone hazard measurements.

[For discussion?] Based on these results, the two data sources were generally
in agreement, but the analysis highlights the difficulty of assessing
storm-related flooding at a multi-county, multi-year scale, as flooding can be
very localized within a county or can \ldots .

However, these frequency maps, together with evidence from
specific tropical cyclones (Figures~\ref{fig:ivanexposure}
and~\ref{fig:jaccard}),  do illustrate the potential for strong differences in
spatial patterns in tropical cyclone exposures, depending on which tropical
cyclone hazards are considered.  A few previous studies have sought to
determine county-level exposure to tropical cyclones over multi-year periods,
including~\textcite{zandbergen2009}, which estimated exposure in \ac{US}
counties to all \ac{US} landfalling Atlantic-basin tropical cyclones
between~1851 and~2003, using both a distance-based metric and a metric that
combined distance and windspeed, and~\textcite{kruk2010}, which explored
exposure to hurricane-related winds in the \ac{US}, including inland areas,
for~1900\,--\,2008.  Our results suggest that such exposure assessments may
perform well in capturing some tropical cyclone hazards (e.g., wind), but
likely miss other potentially dangerous tropical cyclone exposures, especially
for hazards that repeatedly threaten northern or inland counties (e.g., rain,
flooding).

For example, distance from a tropical cyclone's track is relatively easy to
measure and has been used as an operational metric of exposure to tropical
cyclones in previous large-scale studies (examples include ...). Since distance
itself does not constitute a hazard, distance is meant in these cases as a
surrogate to capture exposure to hazards from the tropical cyclone. However,
here we found that in assessing \ac{US} county-level exposure to tropical
cyclones, distance is, at best, a moderate, and often a very poor, surrogate
for exposure to the specific tropical cyclone hazards of high wind, extreme
rainfall, flooding, and tornadoes (Figure~\ref{fig:jaccard}). Therefore, use of
distance to assess tropical cyclone exposure for impact studies could result in
problematic exposure misclassification, which could mask true associations,
even strong associations, between tropical cyclone exposure and outcomes of
interest in impact studies~\parencite{savitz2016interpreting,
armstrong1998effect}.

The use of a single hazard-based metric (e.g., wind) could cause similar
problems if the impact is driven, at least in part, by a different hazard or by
multiple hazards of the tropical cyclone.

In creating this hurricane exposure dataset, we aimed for data that are
available for all eastern \ac{US} counties.

($<$90\% of counties classified into the same wind categories by the two
sources of wind estimates)

The wind radii, unlike the modeled wind speeds, are based on a storm-specific
re-analysis of available data. They therefore have the advantage that they can
capture unusual patterns in specific storms; their main disadvantage is that
they provide categorical, rather than continuous, estimates of wind speed, and
in many cases researchers might prefer more precise estimates of storm wind
speeds. However, since the wind radii data may be useful in some cases as
either a primary metric or for sensitivity analysis, we have included
county-level wind estimates for each storm based on these wind radii, in
addition to the continuous modeled windspeeds, in the open-source data
\parencite{hurricaneexposuredata}.

between county-level classifications of tropical cyclone exposure was typically
low to moderate across the  five exposure metrics we considered

If measurements of one storm hazard are used as a proxy to capture exposure to
a different hazard, or a non-hazard proxy like distance to the storm track is
used to assess exposure to storm hazards, the resulting exposure
misclassification could be differential (i.e., associated with the outcome of
interest or with factors associated with risk of the outcome of interest).

### 2019-10-09 writing/darlings_results.tex
Work on paper, use 34 knot threshold

\section*{Results}

\subsection*{Exposure assessment data, data validation, and software}

\paragraph{Open-source dataset}

We created and published open source data with multi-year, county-level
tropical cyclone exposure data for eastern \ac{US}
counties~\parencite{hurricaneexposuredata}. These data include continuous
county-level measurements for the closest distance of each storm, cumulative
rainfall, and maximum sustained surface wind speed. These continuous metrics
can be used for classifying counties as exposed or unexposed, using thresholds
selected by the user, or can be used as continuous metrics. The dataset also
include binary data on flood and tornado events associated with each storm in
each county. With these data, we published software tools to explore and map
the data and to integrate it with human health
datasets~\parencite{hurricaneexposure}.

We explored potential limitations in these data by comparing them with data
from other available sources.

\paragraph{Rainfall data validation}

For estimates of storm-associated rainfall, we compared estimates from NLDAS-2
re-analysis precipitation data (included as the rainfall estimates in the
open-source data) with ground-based observations in nine sample counties
(Figure~\ref{fig:raincomparison}). Within these counties, storm-related
rainfall measurements were well-correlated between the two data sources, with
rank correlations (bottom right of each graph in
Figure~\ref{fig:raincomparison}) between 0.87 and 0.98. There was some evidence
that the reanalysis data may tend to underestimate rainfall totals in storms
with extremely high rainfall, based a few heavy-rainfall storms in Harris
County, TX, Mobile County, AL, Charleston County, SC, and Wake County, NC
(Figure~\ref{fig:raincomparison}). However, even in these cases, it was rare
for a storm to be classified differently (based on the precipitation threshold
of 75~\si{\milli\metre} we used for exposure classification for later analysis)
based on the source of precipitation data---horizontal and vertical lines in
each small plot in Figure~\ref{fig:raincomparison} show the threshold of
75~\si{\milli\metre}, so storms in the lower left and upper right quadrants
would be classified the same (``exposed'' or ``unexposed'') regardless of the
precipitation data source, while storms in the upper left and lower right
quadrants would be classified differently.

\paragraph{Wind speed data validation}

For maximum sustained surface wind speed estimates, we found that the modeled
estimates, which we included as the primary wind exposure metric in the open
source data and used for further analysis in this study, generally agreed well
with estimates based on the wind radii reported in \ac{HURDAT2}. For most
storms, $\ge$95\% of counties were assigned the same category of wind speed
($<$34~\si{\knot}; 34\,--\,49.9~\si{\knot}; 50\,--\,63.9~\si{\knot};
$\ge$64~\si{\knot}) by both data sources (Figure~\ref{fig:windcomparison}).
Disagreement was limited to large storms that were characterized by high winds
well inland, like Hurricanes Sandy in 2012 and Ike in 2008. For these two
storms, the modeled wind speed values somewhat overestimated the extent of the
storm's most severe winds at landfall but then underestimated the extent of
34\,--\,49.9 \si{\knot} storm winds further inland (Figure~S2). For researchers
who would like to conduct sensitivity analysis between these two sets of wind
data, we have included estimates from the \ac{HURDAT2} wind radii as a
secondary measure of wind speeds in the open-source dataset
\parencite{hurricaneexposuredata}.

\paragraph{Flood data analysis}

For flood data, we compared flood status, based on event listings in the NOAA
Storm Events database, to streamflow measurements at \ac{US} Geological Survey
gages within nine sample counties (Figure~\ref{fig:floodcomparison}).

For flood data, we compared flood status, based on event listings in the NOAA
Storm Events database, to streamflow measurements at \ac{US} Geological Survey
gages within nine sample counties (Figure~\ref{fig:floodcomparison}).  Across
the sampled counties, streamgage data generally indicated higher discharge in
on dates identified as flood events, particularly in counties with more
reporting stream gages (e.g., Harris County, TX, with~10 reporting stream
gages; Figure~\ref{fig:floodcomparison}). There were some cases, however, where
the two flooding data sources were somewhat inconsistent.  For example, there
were one or two tropical cyclones in several of the counties (Mobile County,
AL, Escambia County, FL, and Montgomery County, PA) which were classified as
flood events based on NOAA storm event listings but for which the total
discharge across county streamflow gages was below the 25\textsuperscript{th}
percentile of measurements taken during storms that did not have a flood event
listing. For tropical cyclones that did not have an associated flood listing in
the NOAA Storm Events database, most had lower total streamflow discharge
across county gages compared to storms with flood event listings, and in all
cases but one all gages were below the flooding threshold.  The exception was
one storm in Fulton County, GA (Ida in 2009). This storm occurred in November
2009, following a month with historic rainfall and flooding in Georgia [ref].
Therefore, in this case the flooding associated with Ida was incorporated into
an ongoing flood event listing, with a start date well before the five-day
window used to temporally match storm event listings with tropical cyclone
tracks used here.

\subsection*{Patterns in tropical cyclone exposures in eastern \ac{US} counties}

We used the storm exposure metrics to classify counties as exposed or unexposed
to specific tropical cyclones, and we explored patterns in these exposure
classifications over years with available data
(Table~\ref{tab:exposuresummaries}). Across the five exposure metrics, there
was wide variation in the average number of county exposures per year. For the
tornado metric, there were approximately~50 county exposures per
year, on average, within our study.  For flood and wind metrics, there were
substantially more county exposures ($>$200 / yr on average), and
even more when assessing exposure using the distance and rain metrics ($>$400 /
yr on average).

We also measured the average number of county exposures per storm under each
exposure metric (limiting this analysis to the tropical cyclones that exposed
at least one county to that hazard; Table~\ref{tab:exposuresummaries}).
Tropical cyclones typically exposed only a few counties, if any, to tornadoes
(median counties exposed per storm:~7), more counties under the flood and wind
metrics (median:~26 and~34, respectively), and even more counties for the
metrics of rain (median:~68) and distance (median:~84).  For every metric
except the tornado-based metric, we identified at least one tropical cyclone
with over~300 counties exposed.  However, the largest-extent tropical cyclone
varied by metric (e.g., Beryl in~1994 exposed the most counties based on
distance, Frances in~2004 based on rain, and Ike in~2008 based on wind).

When we calculated and mapped the average number of exposures per decade in
each county for each exposure metric (Figure~\ref{fig:averageexposure}), strong
geographical patterns were clear. Wind-based exposure had a strong coastal
pattern, with almost all exposures in counties within about~200
\si{\kilo\metre} (124~mi) of the coastline. While exposures by rain- and
distance-based metrics were also more common in coastal areas compared to
inland areas, they also created inland exposures that were not identified by
the wind-based metric. Rain and, to some extent, flood exposures were
characterized by a pattern defined by the Appalachian Mountains, with much
lower exposures west of the mountain range than to the east. Almost all
tornado-based exposures were in coastal states, with most in Florida, Alabama,
South Carolina, and North Carolina and almost none north of Maryland.
Flood-based exposures were highest in North Carolina and along the South
Carolina coast, as well as further north in New Jersey, southeastern
Pennsylvania, and southeastern New York. Notably, some areas of the \ac{US}
that experienced regular exposures to tropical cyclone-related excessive
rainfall and floods were not identified as areas of regular risk based on the
more commonly used tropical cyclone exposure assessment metrics of wind or
distance (Figure~\ref{fig:averageexposure}). Patterns were similar when
analysis was restricted to years with all exposure data available
(1996\,--\,2011; Figure~S3).

\subsection*{Agreement in county-specific exposure classification between exposure metrics}

Next, we measured the agreement between exposure classifications for each pair
of metrics. We found that these exposure classifications typically did not
strongly agree between pairs of metrics, indicating that the set of counties
identified as exposed based on one metric can have limited overlap with the
set of counties identified as exposed to the storm by another metric.

Figure~\ref{fig:ivanexposure} shows Hurricane Ivan in 2004 as an example. For
the distance-based metric, the counties assessed as exposed followed the
tropical cyclone's track, including counties well inland. For the wind metric,
only counties near two of the tropical cyclone's landfalls were assessed as
exposed. For rain- and flood-based metrics, however, exposure extended to the
left of the track, including counties as far north as New York and Connecticut,
while for the tornado metric, exposed counties tended to be to the right of the
track and included several counties in central North Carolina, South Carolina,
and Georgia that were not identified as exposed to Ivan based on any other
metric. Figure~S4 provides similar maps for four other example tropical
cyclones (selected because they exposed many \ac{US}  counties based on at
least one metric), and Figure~S5 provides four examples of tropical cyclones in
which the disagreement in exposure assessment between rain- and wind-based
metrics was particularly notable.

We drew similar conclusions when we investigated all~46 tropical cyclones
between~1996 and~2011 (the period for which all five metrics were available)
for which~250 or more counties were exposed based on at least one metric
(Figure~\ref{fig:jaccard}). In this figure, each row provides results for a
specific tropical cyclone, and each box in that row shows the measured Jaccard
coefficient between two of the exposure metrics for that storm.  For most
tropical cyclones, distance-based county-level exposure assessment was at best
in moderate agreement with exposure assessments based on the four hazard-based
metrics. For almost all tropical cyclones, Jaccard index values between
distance and any of the four hazard-based metrics were~$\le0.6$ (i.e., out of
the counties assessed as exposed to the tropical cyclone by at least one of the
two metrics considered,~60\si{\percent} or fewer of the counties had the same
exposure assessment under the two metrics). For many tropical cyclones, the
Jaccard index values were~$\le0.2$, suggesting very poor agreement in exposure
assessment between the metrics.  Among the four hazard-based metrics, the
tornado-based metric showed universally poor agreement with other metrics in
county-level classification across all tropical cyclones considered.  For other
pairs of metrics, there were also generally large differences in which counties
were determined to be exposed, with the Jaccard index values~$<0.4$ for most
metric pairs in most tropical cyclones.

\begin{comment}
Storms are displayed within
clusters that have similar patterns in county-level exposure agreement for
metric pairs, based on hierarchical clustering using the complete link
method~\parencite{murtagh2012algorithms} (i.e., storms in the same cluster tend
to have similar patterns for the pairwise strength of agreement among metrics);
columns are also ordered based on hierarchical clustering. 
Maps are
available showing the counties identified as exposed under each of five metrics
for the widest-extent storm in each cluster: Hurricane Ivan in~2004
(Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd in~1999, Lee in~2011, Cindy
in~2005, and Katrina in~2005 (Figure~S4).
\end{comment}

There were, however, a few exceptions\,---\,tropical cyclones in which similar
counties were determined to be exposed to the tropical cyclone for two or more
of the metrics considered.  For Floyd in~1999 (Figure~S4) and Irene in~2011,
for example, county-level classification agreed moderately to well (Jaccard
index of~$\sim 0.5\mbox{\,--\,}0.8$) for all pairs of exposure metrics except
those including the tornado-based metric. For another set of tropical cyclones
(e.g., Lee in~2011 [Figure~S3], Ernesto in~2006, and Bertha in~1996), there was
moderate to good agreement for all pairwise combinations of distance, rain, and
wind, but poor agreement for all other combinations of metrics. However, these
two sets of tropical cyclones represented the minority of all tropical cyclones
considered; for most tropical cyclones, there was low overlap in the counties
determined to be exposed to the tropical cyclones between most pairings of
metrics.

\subsection*{Case study}

Finally, we conducted a case study to investigate how these differences in
exposure assessment across the exposure metrics could affect estimates of
expected physical exposure among a vulnerable subpopulation,
electricity-dependent Medicare beneficiaries.
Figure~\ref{fig:topelecdependexposure} shows the study counties with the
highest expected physical exposures in this subpopulation based on each
exposure metric. A few counties (Miami-Dade County, FL; Harris County, TX) were
ranked in the top three counties for expected physical exposure in this
subpopulation for almost all exposure metrics
(Figure~\ref{fig:topelecdependexposure}). However, there were key differences
across exposure metrics. For example, most of the top counties based on wind-
and tornado-based metrics were in Florida, while over half of the counties with
highest expected physical exposures for the flood-based metric were in
Mid-Atlantic states (PA, NY).

>>>>

Storms are displayed within
clusters that have similar patterns in county-level exposure agreement for
metric pairs, based on hierarchical clustering using the complete link
method~\parencite{murtagh2012algorithms} (i.e., storms in the same cluster tend
to have similar patterns for the pairwise strength of agreement among metrics);
columns are also ordered based on hierarchical clustering. 
Maps are
available showing the counties identified as exposed under each of five metrics
for the widest-extent storm in each cluster: Hurricane Ivan in~2004
(Figure~\ref{fig:ivanexposure}) and Hurricanes Floyd in~1999, Lee in~2011, Cindy
in~2005, and Katrina in~2005 (Figure~S4).
